A note on $\K$ of Kac---Moody groups over a field

### 2017-08-15 paper.tex
Uncomment counterexample + cosmetics

\newtheorem{df}[lemma]{Definition} \Crefname{df}{Definition}{Definitions}
\newtheorem{example}[lemma]{Example} \Crefname{example}{Example}{Examples}

@@

\begin{document}   
\begin{abstract} We describe a method of computing the $\K$-group associated to any Kac---Moody group over a field.
 The key ingredient in our computation is a variant of Smith normal form of an integer matrix.
\end{abstract}

\maketitle
%\section{Introduction}
The aim of this note is to compute the unstable $\K$-group $\K(A, F)$ for arbitrary generalized Cartan matrix $A$ and $F$ a field,
 and thus give a complete answer to the question raised in~\cite{MW}.
Recall from~\cite{MR}, \cite{Ti} that, by definition, the group $\K(A, F)$ is the kernel of the map 
 $\St(A, F) \to G(A, F)$ between the corresponding Steinberg and Kac---Moody groups over $F$.

@@

Let us first recall the precise statement of the result of Morita and Rehmann.
Denote by $L_i$ either the Milnor group $\K(F)$ or the Milnor---Witt group $\Kmw(F)\cong K_2(2, F)$ 
 depending on whether or not there exists $1 \leq k \leq n$ such that the matrix coefficient $a_{ki}$ is odd.
In either case we denote the Steinberg symbols (resp. cocycles) generating the group $L_i$ by $\{x,y\}_i$, $x, y\in F^\times$
 (see~\cite[Proposition~5.5, Lemme~5.6]{Ma} for the definition of Steinberg cocycles and symbols, respectively).
We always use additive notation for abelian groups $\K(F)$ or $\Kmw(F)$.

\begin{thm}[Morita---Rehmann] \label{thm:mr} 
 $\K(A, F)$ is isomorphic to the quotient $L/J$ where $L = L_1 \oplus L_2 \oplus \ldots \oplus L_n$ and 
 $J$ is the subgroup of $L$ generated by relators $\{x, y^{a_{ji}}\}_i - \{x^{a_{ij}}, y\}_j = 0$.
\end{thm}

@@

Of course, the integers $r_i$ can be concretely obtained as the 
  the diagonal coefficients in the Smith normal form of the $\frac{n(n-1)}{2}\times n$ matrix $\widetilde{A}$ given by
\begin{equation} \label{eq:Aijk} \widetilde{A}_{i<j, k} = \left\{\def\arraystretch{1.2}%
  \begin{array}{@{}c@{\quad}l@{}}
     a_{ij} & \text{ for } k = i, \\
    -a_{ji} & \text{ for } k = j, \\
    0 & \text{otherwise.}    
  \end{array}\right.\end{equation}                    
Notice that the integer $s$ from the statement of~\cref{thm-odd} is precisely the number of nonzero $r_i$'s.

@@

In \cite{MW} M.~Westaway even conjectured that the above two properties could be used to determine $p_d$ from the given $q$ and,
 thus, determine the type of summands appearing in $\K(A, F)$
 (Westaway's matrices $\nu$, $\mu$ appearing in the statement of \cite[Conjecture~9.1]{MW} correspond to our $q$ via $\nu = q^T$, $\mu = {q^{-1}}^T$).
Although, as said before, this characterization works for the matrix $q$ obtained via the algorithm of~\cref{lm:normal-form},
 it may not work for $q$ obtained via the classical algorithm for finding SNF, which ignores columns' parities.
Thus, as stated, \cite[Conjecture~9.1]{MW} is false.
\begin{comment}
\begin{example}
To produce a counterexample, start from GCM 
 $A = \left(\begin{smallmatrix}
 2& -3& -2& 0\\
-1& 2& -2& 0\\
-2& -2& 2& -4\\
0& 0& -4& 2\end{smallmatrix}\right)$
  and write SNF decomposition of~\cref{lm:normal-form} for $\widetilde{A}$
\[\widetilde{A} = \left(\begin{smallmatrix}
-1& 3& 0& 0\\
-2& 0& 2& 0\\
0& 0& 0& 0\\
0& -2& 2& 0\\
0& 0& 0& 0\\
0& 0& -4& 4
\end{smallmatrix} \right) = \left(\begin{smallmatrix}
1& 0& 0& 0& 0& 0\\
2& 1& 0& 0& 0& 0\\
0& 0& 0& 0& 1& 0\\
0& 1& 0& 1& 0& 0\\
0& 0& 0& 0& 0& 1\\
0& -2& 1& 0& 0& 0\end{smallmatrix}\right) \cdot \mathrm{diag}(-1,2,4,4) \cdot \left(\begin{smallmatrix}
1& -3& 0& 0\\
0& -3& 1& 0\\
0& -3& 0& 1\\
0& 1& 0& 0\\ \end{smallmatrix}\right) = p \cdot d \cdot q.\]
Looking at the last 2 columns of $q$ we deduce that columns with $1$st and $4$th columns of $d$ are odd while $2$nd and $3$rd are even.
Now take $s = t_{43}(1) \in \E(4, \ZZ)$, $t = t_{43}(-1) \ in \E(6, \ZZ)$.
Notice that the central factor in the decomposition $\widetilde{A} = (p t) \cdot (t^{-1} d s^{-1}) \cdot (s q)$ is still 
 a still diagonal matrix but neither of the above two criterions is applicable to the matrix $sq$.
\end{example}
\end{comment}
\printbibliography

>>>>

\newtheorem{df}[lemma]{Definition} \Crefname{df}{Definition}{Definitions}
\newtheorem{example}{Example} \Crefname{example}{Example}{Examples}

@@

\begin{document}   
\begin{abstract} We describe a method of computing the $\K$-group associated to any Kac--Moody group over a field.
 The key ingredient in our computation is a variant of Smith normal form of an integer matrix.
\end{abstract}

\maketitle
%\section{Introduction}
The aim of this note is to compute the unstable $\K$-group $\K(A, F)$ for arbitrary generalized Cartan matrix $A$ and $F$ a field,
 and thus give a complete answer to the question raised in~\cite{MW}.
Recall from~\cite{MR}, \cite{Ti} that, by definition, the group $\K(A, F)$ is the kernel of the map 
 $\St(A, F) \to G(A, F)$ between the corresponding Steinberg and Kac--Moody groups over $F$.

@@

Let us first recall the precise statement of the result of Morita and Rehmann.
Denote by $L_i$ either the Milnor group $\K(F)$ or the Milnor--Witt group $\Kmw(F)\cong K_2(2, F)$ 
 depending on whether or not there exists $1 \leq k \leq n$ such that the matrix coefficient $a_{ki}$ is odd.
In either case we denote the Steinberg symbols (resp. cocycles) generating the group $L_i$ by $\{x,y\}_i$, $x, y\in F^\times$
 (see~\cite[Proposition~5.5, Lemme~5.6]{Ma} for the definition of Steinberg cocycles and symbols, respectively).
We always use additive notation for abelian groups $\K(F)$ or $\Kmw(F)$.

\begin{thm}[Morita--Rehmann] \label{thm:mr} 
 $\K(A, F)$ is isomorphic to the quotient $L/J$ where $L = L_1 \oplus L_2 \oplus \ldots \oplus L_n$ and 
 $J$ is the subgroup of $L$ generated by relators $\{x, y^{a_{ji}}\}_i - \{x^{a_{ij}}, y\}_j = 0$.
\end{thm}

@@

Of course, the integers $r_i$ can be concretely obtained as
  the diagonal coefficients in the Smith normal form of the $\frac{n(n-1)}{2}\times n$ matrix $\widetilde{A}$ given by
\begin{equation} \label{eq:Aijk} \widetilde{A}_{i<j, k} = \left\{\def\arraystretch{1.2}%
  \begin{array}{@{}c@{\quad}l@{}}
     a_{ij} & \text{ for } k = i, \\
    -a_{ji} & \text{ for } k = j, \\
    0 & \text{otherwise.}    
  \end{array}\right.\end{equation}                    
Notice that the integer $s$ from the statement of~\cref{thm-odd} is precisely the number of nonzero $r_i$'s.

@@

In \cite{MW} M.~Westaway even conjectured that the above two properties could be used to determine $p_d$ from the given $q$ and,
 thus, determine the type of summands appearing in $\K(A, F)$
 (Westaway's matrices $\nu$, $\mu$ appearing in the statement of \cite[Conjecture~9.1]{MW} correspond to our $q$ via $\nu = q^T$, $\mu = {q^{-1}}^T$).
Although, as said before, this characterization works for the matrix $q$ obtained via the algorithm of~\cref{lm:normal-form},
 it may not work for $q$ obtained via the classical algorithm for finding SNF, which ignores columns' parities.
Thus, as stated, \cite[Conjecture~9.1]{MW} is false.

\begin{example}
To produce a concrete counterexample, start with GCM 
 $A = \left(\begin{smallmatrix}
 2& -3& -2& 0\\
-1& 2& -2& 0\\
-2& -2& 2& -4\\
0& 0& -4& 2\end{smallmatrix}\right)$
  and write SNF decomposition of~\cref{lm:normal-form} for $\widetilde{A}$
\[\widetilde{A} = \left(\begin{smallmatrix}
-1& 3& 0& 0\\
-2& 0& 2& 0\\
0& 0& 0& 0\\
0& -2& 2& 0\\
0& 0& 0& 0\\
0& 0& -4& 4
\end{smallmatrix} \right) = \left(\begin{smallmatrix}
1& 0& 0& 0& 0& 0\\
2& 1& 0& 0& 0& 0\\
0& 0& 0& 0& 1& 0\\
0& 1& 0& 1& 0& 0\\
0& 0& 0& 0& 0& 1\\
0& -2& 1& 0& 0& 0\end{smallmatrix}\right) \cdot \mathrm{diag}(-1,2,4,4) \cdot \left(\begin{smallmatrix}
1& -3& 0& 0\\
0& -3& 1& 0\\
0& -3& 0& 1\\
0& 1& 0& 0\\ \end{smallmatrix}\right) = p \cdot d \cdot q.\]
Looking at the last 2 columns of $q$ we deduce that $1$st and $4$th columns of $d$ are odd while $2$nd and $3$rd ones are even.
Now take $s = t_{43}(1) \in \SL(4, \ZZ)$, $t = t_{43}(-1) \in \SL(6, \ZZ)$.
Notice that the central factor in the decomposition $\widetilde{A} = (p t) \cdot (t^{-1} d s^{-1}) \cdot (s q)$ is still 
 a diagonal matrix but neither of the above two criterions is applicable to the matrix $sq$.
\end{example}

\printbibliography
